Transition path sampling based calculations of free energies for enzymatic reactions: the case of human methionine adenosyl transferase and plasmodium vivax adenosine deaminase

### 2022-06-10 tex/resubmission1/response/response.tex

%\begin{response}
%{The current work shows the application of the BOLAS method to two enzyme-catalyzed reactions. 
%The BOLAS method is a transition path sampling algorithm that can be extended to calculate free 
%energies along a pre-defined reaction coordinate. There are already several methods to calculate 
%free energy profiles and many have been applied to enzymatic reactions, but none is a black-box, 
%and therefore the implementation and application of this method to an enzymatic reaction is an 
%important advance. The text is well written and clear.}
%reply
%TPS based free energy calculations have been attempted previously for the calculations
%involving model problems involving analytical potential energy surfaces or for 
%configurational or structural changes of an enzyme. A chemical reaction catalyzed by 
%an enzyme can involve a complicated reaction coordinate and TPS based method using the 
%shooting algorithm provides a bias free technique for calculating the free energy
%of the reaction.

%\end{response}

\begin{response}
{First one needs to know how the specific parameters of the method affect 
the results. Or at least, how or why they were chosen. If not for all parameters, 
at least for the most relevant ones (but this needs also to be specified). 
For example:
* why use 150 reactive trajectories?}
%reply

%The application of TPS for stuyding the kinetics and transition states of 
%enzyme catalyzed reactions has been repeorted several times by our group as well
%as by others we decided to not list out every detail. But we have now included 
%some details about the choice the number of trajectories and so on.
The number of reactive trajectories in a TPS ensemble is decided based 
on how quickly the shooting algorithm results in decorrelation of the reactive 
trajectories from the initial biased trajectory. If the decorrelation is fast 
then fewer trajectoreis are sufficient for further committor analysis. 
To be sure that we do not use trajecotries that are correlated to the initial 
biased trajectory we ignore the first 100 trajectories for committor analysis. 
Additionally we have added a discussion on the acceptance percentage of reactive 
trajectories in the manuscript. It has been observed that when the acceptance 
percentage is low the trajectories decorrelate faster.  
\end{response}

\begin{response}
  {why should each reaction last 0.5 ps? How is it known if that time is too short 
  or too long?}
We used longer trajectories (1 ps) during the initial stages of our calculations
and checked to make sure that the reaction occurs within
the 0.5 ps time length before collecting the TPS ensemble. 
\end{response}

@@

{\begin{response}
{* why 20 windows?}
%reply
This choice was made based on the order parameter window that we are scanning.
20 windows results in a window size of 0.39 angstroms and we use a overlap of
0.09 angstroms on either side of the window to make sure that we have configurations
that are common to neighboring windows which is necessary to obtain a continuous free energy
profile. Smaller window sizes result in inefficient sampling because the TPS based method
does not include bias forces to keep the trajectories within the windows and we end up
sampling outside the windows especially in regions where the free energy has a steeper slope.  
\end{response}

\begin{response}
{* When calculating the commitor, you cannot say "repeating this procedure many times". 
How many are needed to obtain a given precision or to obtain the given figure?
\newline
* As before, when calculating the separatrix, you can not say "repeating this 
process for several trajectories".}
%reply
Since this is the discussion of the committor analysis and committor distribution 
analysis in general and not for the specific case for the enzymes, we had not
provided the number of trajectories that were used in these calculations. 
\end{response}

@@

\begin{response}{Beyond these general issues raised so far, there are some other points 
that deserve some attention.
\newline
1. How is the initial reactive trajectory generated? Although obtaining an initial reactive 
trajectory is not the crux of the algorithm it is a non-trivial step that should be more 
detailed. In particular, for the pvADA enzyme, which involves a two-step reaction, how 
do the way these 2 steps take place in the initial trajectory condition of the TPS ensemble?}

% REPLY
The reaction catalyzed by the \textit{pv}ADA enzyme consists of 
the nucleophilic attack of the OH$^{-}$ on the C6 atom, proton transfer from the 
GLU 223 residue to the N1 atom in the adenosyl moeity and finally the 
the transfer of the proton from the OH$^{-}$ to the N6 atom resulting in the formation 
of NH$_3$ and inosine. We have included the time series of the key distances in 
the supporting information which shows that we collect trajectories which contain all these
steps taking the system from the reactant state to the product state with the order of events
as predicted from experiments. 
\end{response}

\begin{response}{2. Related to the previous question, considering that the barriers 
surrounding the intermediate are higher than kT, shouldn't the trajectories remain 
trapped in the intermediate? In other words, isn't the duration of the whole trajectory 
determining the mechanism by enforcing a ballistic trajectory? Is the d(NC)-d(OC) still 
a good reaction coordinate to describe the second step? If not, shouldn't that underestimate 
the barrier, something that should be compensated by a (not calculated) transmission coefficient?}

% REPLY
The calculated geometry of the intermediate resembles that of the Meissenheimer complex which is
similar to that predicted from experiments. In our TPS reactive trajectories we observe that 
once the complex is formed there is a rapid proton transfer from the nucleophilic OH$^{-}$ 
to the NH$_2$ group resulting in the release of NH$_3$ and the free energy change due to the 
proton transfer is 1.7 kCal mol$^{-1}$. The order parameter that is chosen is based on the fact that
it allows for the best definition of the reactant and product states.    
\end{response}

\begin{response}
{3. I find the free energy profiles in fig 7 and 11 worrying. Take the 4th window in 
Fig. 7: it is clearly showing a smaller slope than the interpolation and this is because 
there are jumps between the windows. The first point of this window and the next one are 
"disconnected" from the other ones. Why is this so? Repeating the calculations with smaller 
windows would show the effect of window size on the result (that would address one of my 
questions regarding the choice of parameters).}
% REPLY
Since there are no bias forces constraining the trajectories to stay within a defined window, 
in regions where the free energy curve has a large slope the trajectories tend to only sample 
configurations almost completely outside the window with acceptance percentages $\leq 10$.
This disconnection is simply an artifact of the histogram 
\end{response}

### 2022-06-13 tex/resubmission1/response/response.tex

We would like to thank the Reviewer for his/her time and for evaluating
the manuscript.

%\begin{response}
%{The current work shows the application of the BOLAS method to two enzyme-catalyzed reactions. 
%The BOLAS method is a transition path sampling algorithm that can be extended to calculate free 
%energies along a pre-defined reaction coordinate. There are already several methods to calculate 
%free energy profiles and many have been applied to enzymatic reactions, but none is a black-box, 
%and therefore the implementation and application of this method to an enzymatic reaction is an 
%important advance. The text is well written and clear.}
%reply
%TPS based free energy calculations have been attempted previously for the calculations
%involving model problems involving analytical potential energy surfaces or for 
%configurational or structural changes of an enzyme. A chemical reaction catalyzed by 
%an enzyme can involve a complicated reaction coordinate and TPS based method using the 
%shooting algorithm provides a bias free technique for calculating the free energy
%of the reaction.

%\end{response}

\begin{response}
{First one needs to know how the specific parameters of the method affect 
the results. Or at least, how or why they were chosen. If not for all parameters, 
at least for the most relevant ones (but this needs also to be specified). 
\newline
For example:
* why use 150 reactive trajectories?}
%reply

%The application of TPS for stuyding the kinetics and transition states of 
%enzyme catalyzed reactions has been repeorted several times by our group as well
%as by others we decided to not list out every detail. But we have now included 
%some details about the choice the number of trajectories and so on.
The number of reactive trajectories in a TPS ensemble is decided based 
on how quickly the shooting algorithm results in decorrelation of the reactive 
trajectories from the initial biased trajectory. If the decorrelation is fast 
then fewer trajectoreis are sufficient for further committor analysis. 
To be sure that we do not use trajecotries that are correlated to the initial 
biased trajectory we ignore the first 100 trajectories for committor analysis. 
Additionally we have added a discussion on the acceptance percentage of reactive 
trajectories in the manuscript. 
%It has been observed that when the acceptance 
%percentage is low the trajectories decorrelate faster.  
\end{response}

@@

\begin{response}
  {why should each reaction last 0.5 ps? How is it known if that time is too short 
  or too long?}
We used longer trajectories (1 ps) during the initial stages of our calculations
and checked to make sure that the reaction occurs within
the 0.5 ps time length before collecting the TPS ensemble. 
\end{response}

{\begin{response}
{* why 20 windows?}
%reply
This choice was made based on the order parameter window that we are scanning.
20 windows results in a window size of 0.39 angstroms and we use a overlap of
0.09 angstroms on either side of the window to make sure that we have configurations
that are common to neighboring windows which is necessary to obtain a continuous free energy
profile. Smaller window sizes result in inefficient sampling because the TPS based method
does not include bias forces to keep the trajectories within the windows and we end up
sampling outside the windows especially in regions where the free energy has a steeper slope.  
\end{response}

\begin{response}
{* When calculating the commitor, you cannot say "repeating this procedure many times". 
How many are needed to obtain a given precision or to obtain the given figure?
\newline
* As before, when calculating the separatrix, you can not say "repeating this 
process for several trajectories".}
%reply
We have included a few lines in the committor analysis 
discussion where the details of the number of trajectories used is mentioned.  
\end{response}

\begin{response}{Some of these questions arise because nowhere in 
the manuscript are confidence intervals given. Neither in the energy 
profile nor in the transition state structure distances (Fig. 5). 
Therefore the effect of these parameters on the precision of the 
results is not known.
}
We have added a supporting information file which contains geometrical 
parameters of the transition state structures that we had calculated 
along with the mean and standard deviations. 
The transition state structures depicted in Figs. 5 and 10 are 
representatives from the ensemble. We have also updated Figs. 7 
and 11 to contain standard deviations of the free energy that were
calculated using the bootstrapping method within the windows.  
\end{response}

@@

\begin{response}{Some of these questions also arise because the 
code to implement the method is not made public. Doing so would 
not only clarify some questions but also prevent others from 
reinventing the wheel, that is, coding the implementation again 
to check the results of this paper or to apply the method to their specific enzymes.}

% REPLY
We agree that the availability of code is necessary for reproducibility of the results.
TPS based free energies calculations are akin to vanilla TPS calculations with 
only a few modifications to the acceptance criteria as discussed in the manuscript as well
as in the original BOLAS paper. We followed the algorithm stated in the manuscript
and implemented it as python scripts but it was hard coded to model the MAT2A and 
$\textit{pv}$ADA catalyzed reactions. We would release it in a public repository once it is 
tailored for general purpose usage.  
\end{response}

\begin{response}{The second piece of information needed is how the method 
performs compared to others. Both in terms of computational cost and in 
terms of accuracy. Comparison with experiment is interesting but not very 
enlightening as the computational results depend on the method to calculate 
the free energy profile (tested here) but also on the chosen reaction coordinate, 
the QM region and the QM method, among others. A comparison with Umbrella Sampling, 
which also uses a reaction coordinate and also uses windows would be ideal. Again, 
to make fair comparisons we would need error bars to estimate the convergence of 
the results and reports of computational costs.}

We have added discussions in the manuscript and a supporting information 
document on the efficiency of sampling, 
statistical analysis of the transition state structures and free energy profiles. 
% REPLY
\end{response}

@@

\begin{response}{Beyond these general issues raised so far, there are some other points 
that deserve some attention.
\newline
1. How is the initial reactive trajectory generated? Although obtaining an initial reactive 
trajectory is not the crux of the algorithm it is a non-trivial step that should be more 
detailed. In particular, for the pvADA enzyme, which involves a two-step reaction, how 
do the way these 2 steps take place in the initial trajectory condition of the TPS ensemble?}

% REPLY
The reaction catalyzed by the \textit{pv}ADA enzyme consists of 
the nucleophilic attack of the OH$^{-}$ on the C6 atom, proton transfer from the 
GLU 223 residue to the N1 atom in the adenosyl moeity and finally the 
the transfer of the proton from the OH$^{-}$ to the N6 atom resulting in the formation 
of NH$_3$ and inosine. We have included the time series of the key distances in 
the supporting information which shows that we collect trajectories which contain all these
steps taking the system from the reactant state to the product state with the order of events
as predicted from experiments. 
\end{response}

\begin{response}{2. Related to the previous question, considering that the barriers 
surrounding the intermediate are higher than kT, shouldn't the trajectories remain 
trapped in the intermediate? In other words, isn't the duration of the whole trajectory 
determining the mechanism by enforcing a ballistic trajectory? Is the d(NC)-d(OC) still 
a good reaction coordinate to describe the second step? If not, shouldn't that underestimate 
the barrier, something that should be compensated by a (not calculated) transmission coefficient?}

% REPLY
The calculated geometry of the intermediate resembles that of the Meissenheimer complex which is
similar to that predicted from experiments. In our TPS reactive trajectories we observe that 
once the complex is formed there is a rapid proton transfer from the nucleophilic OH$^{-}$ 
to the NH$_2$ group resulting in the release of NH$_3$ and the free energy change due to the 
proton transfer is 1.7 kCal mol$^{-1}$. 
%The order parameter that is chosen is based on the fact that
%it allows for the best definition of the reactant and product states.    
\end{response}

\begin{response}
{3. I find the free energy profiles in fig 7 and 11 worrying. Take the 4th window in 
Fig. 7: it is clearly showing a smaller slope than the interpolation and this is because 
there are jumps between the windows. The first point of this window and the next one are 
"disconnected" from the other ones. Why is this so? Repeating the calculations with smaller 
windows would show the effect of window size on the result (that would address one of my 
questions regarding the choice of parameters).}
% REPLY
Since there are no bias forces constraining the trajectories to stay within a defined window, 
in regions where the free energy curve has a large slope the trajectories tend to only sample 
configurations that are almost completely outside the window with acceptance percentages $\leq 10$.
This results in populations that are skewed in favor of regions with lower free energies. We sampled 
3000 trajectories within each window and made sure that the change in the free energy profile
is negligible as a function of number of sampled trajectories to make sure we have achieved 
convergence. Moreover the largest standard deviations that we obtained from bootstrapping 
analysis was 0.9 kCal mol$^{-1}$ which also provides a measure for convergence.

@@

\end{response}

\begin{response}
{4. The author should check they cite the appropriate works. I am afraid 
ref 20 is not an application of TPS to enzyme catalysis. To my knowledge, The first 
applications of TPS to enzyme catalysis are  $J. Am. Chem. Soc.2005, 127, 13822-13831$ 
and $J. Phys. Chem. B 2007, 111, 20, 5708-5718$.}
% REPLY

We thank the Reviewer for pointing out this mistake. We have made the appropriate changes to that
citation in the manuscript. 
\end{response}

\begin{response}{Overall this is a relevant work, but the authors need to go 
deeper into the scientific details to be useful to other members of the 
community, because what the manuscript should be describing is not that the 
method works --we already knew that-- but how to make it work in enzymatic reactions.}
% REPLY
\end{response}
We thank the reviewer for the constructive comments to make the manuscript better. We have 
addressed the questions by adding more statistical analysis and discussions of the details 
of the calculations. 
\section*{Other Changes}

### 2022-06-13 tex/resubmission1/manuscript.tex
added new changes

repeating this procedure many times and finding the fraction of trajectories that go 

@@

ensemble results in a collection of transition states called the separatrix.

@@

Fig. \ref{fig:mat2a-fenergy} where the barrier is $16\;\text{Kcal mol}^{-1}$.
This is in excellent agreement with the experimental result.  

@@

@@

The MAT2A enzymatic reaction was analyzed in detail to elucidate the transition state,
reaction coordinate and the free energy profile. The reaction proceeds through a 
S$_\text{N}$2 mechanism, the d$_{\text{OC}}$ and d$_{\text{SC}}$ are $2.09$ {\AA}
and $2.31$ {\AA}, respectively. The reaction coordinate is determined to involve  
the Gln113, Ser114, Arg249 and Arg264 residues. 
The free energy barrier for this reaction was calculated to be 
$16\;\text{Kcal}\;\text{mol}^{-1}$ which is in good agreement 
with the experimentally measured activation energy of $17.27\;\pm\;1.5\;\text{Kcal mol}^{-1}$. 
Furthermore the reaction catalyzed by the $pv$ADA enzyme is analyzed in detail to obtain the  
transition state and calculate the free energy profile.
The transition state has a Meisenheimer complex like structure with the 
d$_\text{NC}$ and d$_\text{OC}$ being $1.62$ {\AA} and $1.49$ {\AA}, respectively.
The free energy barrier for this reaction for found to be $21\;\text{Kcal}\;\text{mol}^{-1}$
with a double well shape similar to that measured from experiments. 
These results show that free energy simulations based on unbiased transition 
path sampling computations can result in accurate and computationally
efficient predictions of both free energy profiles and mechanisms.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}
SGB acknowledges funding from the NIH through grant number GM127594.
\end{acknowledgement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The same is true for Supporting Information, which should use the
%% suppinfo environment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{suppinfo}

%\end{suppinfo}

>>>>

repeating this procedure at least 50 times for each time slice and finding the fraction of trajectories that go 

@@

ensemble results in a collection of transition states called the separatrix. We calculated 
11 transition state structures for the MAT2A and the \textit{pv}ADA enzymes and details 
regarding them are reported in the supporting information. 

@@

Fig. \ref{fig:mat2a-fenergy}. It includes standard deviations calculated 
using the bootstrapping method for the populations of configurations 
sampled within each window. \cite{Hub10JChemTheoryComput6p3713}
The calculated free energy barrier is $16\;\text{Kcal mol}^{-1}$.
which is in excellent agreement with the experimental result.  

@@

The free energy profile along with standard deviations calculated 
using the bootstrapping method for the populations of configurations 
sampled within each window \cite{Hub10JChemTheoryComput6p3713} is 
depicted in Fig. \ref{fig:ada-fenergy}.

@@

The MAT2A enzymatic reaction was analyzed in detail to elucidate the transition state,
reaction coordinate and the free energy profile. The reaction proceeds through a 
S$_\text{N}$2 mechanism, the d$_{\text{OC}}$ and d$_{\text{SC}}$ are $2.09$ {\AA}
and $2.31$ {\AA}, respectively. The reaction coordinate is determined to involve  
the Gln113, Ser114, Arg249 and Arg264 residues. 
The free energy barrier for this reaction was calculated to be 
$16\;\text{Kcal}\;\text{mol}^{-1}$ which is in good agreement 
with the experimentally measured activation energy of $17.27\;\pm\;1.5\;\text{Kcal mol}^{-1}$. 
Furthermore the reaction catalyzed by the $pv$ADA enzyme is analyzed in detail to obtain the  
transition state and calculate the free energy profile.
The transition state has a Meisenheimer complex like structure with the 
d$_\text{NC}$ and d$_\text{OC}$ being $1.62$ {\AA} and $1.49$ {\AA}, respectively.
The free energy barrier for this reaction for found to be $21\;\text{Kcal}\;\text{mol}^{-1}$
with a double well shape similar to that measured from experiments. 
These results show that free energy simulations based on unbiased transition 
path sampling computations can result in accurate and computationally
efficient predictions of both free energy profiles and mechanisms.
Free energy calculations within the TPS shooting algorithm do not employ any 
bias forces constraining the trajectories within a defined window. 
We observe that in regions where the free energy curve has a large 
slope the trajectories tend to majorly sample configurations that are 
outside the window with acceptance percentages $\leq 10$. While this improves
the decorrelation time of the sampled trajectories, techniques within 
TPS to enhance the sampling in these regions will be necessary for its
widespread applications to complex biomolecular systems.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}
SGB acknowledges funding from the NIH through grant number GM127594.
\end{acknowledgement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The same is true for Supporting Information, which should use the
%% suppinfo environment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{suppinfo}
The supporting information is available free of charge at

\end{suppinfo}

### 2022-06-13 tex/resubmission1/supp-info/si.tex
added new figure and some more text
added new changes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{tocentry}
\section{Initial reactive trajectory}
The reaction catalyzed by \textit{pv}ADA is between adenosine and OH$^{-}$ anion resulting
in the formation of inosine. Experimental studies suggest that the GLU 223 residue 
protonates the N1 atom of the adenosyl moeity and this transfer of proton is almost complete 
at the transition state. \cite{Luo07JAmChemSoc129p8008} The protonation of the N1 atom 
as observed in a representative reactive trajectory from the TPS ensemble is shown in 
Fig. \ref{sfig:he2time}.

@@

\begin{figure}
\includegraphics[scale=1]{figures/ada-oh-nh60.pdf}
\caption{The distance between the N6-H1 atoms and O1-H1 atoms as a function of time. The proton transfer from the hydroxide
anion to the NH$_2$ group occurs after the nucleophilic attack of the O1 atom on the C6 atom.}
\end{figure}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{tocentry}
\section{\textit{pv}ADA catalyzed reaction}
The reaction catalyzed by \textit{pv}ADA is between adenosine and OH$^{-}$ anion resulting
in the formation of inosine. Experimental studies suggest that the GLU 223 residue 
protonates the N1 atom of the adenosyl moeity and this transfer of proton is almost complete 
at the transition state. \cite{Luo07JAmChemSoc129p8008} The protonation of the N1 atom 
as observed in a representative reactive trajectory from the TPS ensemble is shown in 
Fig. \ref{sfig:he2time}. The synamics of the proton transfer from the OH group to the 
NH$_2$ group which are both bonded to the adenosine moeity forming a Meissenheimer complex is 
depicted in Fig. \ref{sfig:n6proton}.

@@

\begin{figure}
\includegraphics[scale=1]{figures/ada-oh-nh60.pdf}
\caption{The distance between the N6-H1 atoms and O1-H1 atoms as a function of time. The proton transfer from the hydroxide
anion to the NH$_2$ group occurs after the nucleophilic attack of the O1 atom on the C6 atom.}
\label{sfig:n6proton}
\end{figure}

@@

%\section{Bootstrapping analysis of the free energy curves}
